---
title: "PSTAT131, Homework 2"
author: "Sofia Spasibenko"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE,
                      warning = FALSE)
library(tidyverse)
library(tidymodels)
library(ggplot2)
library(discrim)
library(poissonreg)
library(corrr)
library(yardstick)
```

## Linear Regression

```{r}
abalone = read_csv("/Users/Sofia/Desktop/PSTAT131/homework-2/data/abalone.csv")

```

### Question 1

### Your goal is to predict abalone age, which is calculated as the number of rings plus 1.5. Notice there currently is no `age` variable in the data set. Add `age` to the data set.

```{r}
abalone <- mutate(abalone, age = rings + 1.5)
ggplot(abalone, aes(x=age))+ geom_bar(stat = 'count', fill = 'blue')
```

### Assess and describe the distribution of `age`.
The distribution of age seems to be a right-skewed Normal curve centered around 10.

### Question 2

### Split the abalone data into a training set and a testing set. Use stratified sampling. You should decide on appropriate percentages for splitting the data.

```{r}
set.seed(128)

abalone_split <- initial_split(abalone, prop = 0.70,
                                strata = age)
abalone_tr <- training(abalone_split)
abalone_te <- testing(abalone_split)
```

*Remember that you'll need to set a seed at the beginning of the document to reproduce your results.*

### Question 3

###Using the **training** data, create a recipe predicting the outcome variable, `age`, with all other predictor variables. Note that you should not include `rings` to predict `age`. Explain why you shouldn't use `rings` to predict `age`.

We shouldn't use rings to predict age since we are trying to predict their age by using their outside appearance.

Steps for your recipe:

1.  dummy code any categorical predictors

2.  create interactions between

    -   `type` and `shucked_weight`,
    -   `longest_shell` and `diameter`,
    -   `shucked_weight` and `shell_weight`

3.  center all predictors, and

4.  scale all predictors.

You'll need to investigate the `tidymodels` documentation to find the appropriate step functions to use.

```{r}
abalone_recipe <- recipe(age ~ type + longest_shell + diameter + height +
                           whole_weight + shucked_weight + viscera_weight +
                           shell_weight, data = abalone_tr) %>%
  step_dummy(type) %>%  #dummy code for type in abalone
  step_interact(terms = ~ shucked_weight:starts_with("type_") ) %>%  #interactions
  step_interact(terms = ~ longest_shell:diameter) %>%
  step_interact(terms = ~ shucked_weight:shell_weight) %>%
  step_center(all_numeric_predictors()) %>%  #center all predictors
  step_scale(all_numeric_predictors()) %>% #scale all predictors
  prep()

bake(abalone_recipe, new_data = NULL)

# abalone_mod <- abalone_recipe 
```

### Question 4

Create and store a linear regression object using the `"lm"` engine.

```{r}
lm_model <- linear_reg() %>% 
  set_engine("lm")
```


### Question 5

Now:

1.  set up an empty workflow,
2.  add the model you created in Question 4, and
3.  add the recipe that you created in Question 3.

```{r}
lm_wflow <- workflow() %>%    #empty workflow
  add_model(lm_model) %>%     #model from Question 4
  add_recipe(abalone_recipe)  #recipe from Question 3
```


### Question 6

Use your `fit()` object to predict the age of a hypothetical female abalone with longest_shell = 0.50, diameter = 0.10, height = 0.30, whole_weight = 4, shucked_weight = 1, viscera_weight = 2, shell_weight = 1.

```{r}
lm_fit <- fit(lm_wflow, abalone_tr)
new <- data.frame(type = 'F', longest_shell = 0.50, diameter = 0.10, 
                  height = 0.30, whole_weight = 4, shucked_weight = 1,
                  viscera_weight = 2, shell_weight = 1)
predict(lm_fit, new_data = new)
```
Our model predicts 23.29291 as the age value.

### Question 7

Now you want to assess your model's performance. To do this, use the `yardstick` package:

1.  Create a metric set that includes *R^2^*, RMSE (root mean squared error), and MAE (mean absolute error).
2.  Use `predict()` and `bind_cols()` to create a tibble of your model's predicted values from the **training data** along with the actual observed ages (these are needed to assess your model's performance).
3.  Finally, apply your metric set to the tibble, report the results, and interpret the *R^2^* value.

```{r}
multi_metric <- metric_set(rsq, rmse, mae)  #1. make the metric set
abalone_tr_res <- predict(lm_fit, 
                          new_data = abalone_tr %>% select(-(rings:age))) #2. making tibble
abalone_tr_res <- bind_cols(abalone_tr_res, abalone_tr %>% select(age))
abalone_tr_res %>% 
  head()
multi_metric(abalone_tr_res, truth = age,  #3. applying metric set
                estimate = .pred)
```

Our model performance returns a 0.5603852	*R^2^* value, a 2.1294926	root mean squared error value, and a 1.5328272	mean absolute error value. The *R^2^* value tells us that about 56% of variability in age can be explained by the predictors, which isn't particularly high, confirming that abalone ages are hard to predict.
